---
title: "PS6_Rodrigo_Valdes"
author: "Rodrigo Valdes"
date: "February 20, 2017"
output:
    github_document:
      toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/Users/ruy/Documents/UChicago/Winter_2017/pcm/persp-model/students/valdesortiz_rodrigo/ps6/')
library(tidyverse)
library(modelr)
library(pROC)
library(modeest)
library(broom)
library(forcats)
data1 <- read_csv('data/mental_health.csv')
data2 <- read_csv('data/gss2006.csv')
```

## Describe the data

1.1. Histogram

```{r}
data1 %>%
  group_by(vote96) %>%
  count() %>%
  ggplot(aes(x=as.character(vote96), y=n/sum(n))) +
  geom_bar(stat='identity') +
  labs(title = "Turnout distribution",
       subtitle = '0 = No Voted , 1 = Voted, NA = missing value',
       x = 'Voter turnout',
       y = 'Fraction of voters')
```

To calculate the conditional probability, I will only use the data for the people that we know if voted or not. That is to say; I will not consider the missing values. It can produce a small bias, but it is the best assumption given the structure of the data.
```{r}
# unconditional probability
prob <- data1 %>%
  group_by(vote96) %>%
  filter(vote96!='NA') %>%
  count() %>%
  mutate('unconditional probability' = n/sum(n)) %>%
  select(-n)

```

Unconditional probabilities.
The zero represents not vote.

```{r}
prob
```

1.2. Plot

According to the graph, depressed people tend to vote less. The correlation between mental health and proportion of vote turnout is negative. Then, there is an insight that the theory that says that depressed people tend to participate less in elections is correct. 

```{r}
# scatterplot
data1 %>%
  ggplot(aes(mhealth_sum, vote96)) +
  geom_jitter(na.rm=TRUE, height=.03, width=.3, alpha=.3) +
  geom_smooth(method='lm') +
  labs(title = "Voter turnout by mental health state",
       subtitle = '0 = no , 1 = yes; missing values are removed',
       x = 'Mental health ',
       y = 'Proportion of vote turnout')
```
Note: I utilize jitter to show there are many observations by mental health index.

## Basic Model

# BM - 1

The summary of the model depicts that the coefficient for mental health is highly significant, well above the 1%. For the case of mental health equal two, the median, the change in the probability of vote when the index in mental health increases from two to three is: -0.031. Furthermore, the same relationship for a change from ten to eleven is -0.035. Then, the number is substantial, especially regarding the change from completely mental health to completely depressed. For instance, the change in the probability of vote for a healthy person (index zero) against an entirely depressed individual (index fifteen) is about 0.5. Then, the mental health can affect the probability of vote drastically.  

```{r}
fit_logistic_sim <- glm(vote96 ~ mhealth_sum, data=data1, family=binomial())
summary(fit_logistic_sim)

# mlv(data1$mhealth_sum, method = "mfv", na.rm=TRUE)
# mean(data1$mhealth_sum, na.rm=TRUE)
median(data1$mhealth_sum, na.rm=TRUE)

# Evaluation on changes in mental health

exponential_part_3 <- exp(1.13921 + -0.14348 * 3)
exponential_part_2 <- exp(1.13921 + -0.14348 * 2)

change <- ((exponential_part_3)/(1+exponential_part_3) - (exponential_part_2)/(1+exponential_part_2))
change

exponential_part_11 <- exp(1.13921 + -0.14348 * 11)
exponential_part_10 <- exp(1.13921 + -0.14348 * 10)
change2 <- ((exponential_part_11)/(1+exponential_part_11) - (exponential_part_10)/(1+exponential_part_10))
change2

exponential_part_15 <- exp(1.13921 + -0.14348 * 15)
exponential_part_0 <- exp(1.13921 + -0.14348 * 0)
change3 <- ((exponential_part_15)/(1+exponential_part_15) - (exponential_part_0)/(1+exponential_part_0))
change3

```

# BM -  2. 

The coefficient for mental health, -0.14348, is the change in log-odds when mental health changes in one unit. The graph shows the linear relationship in the log-odds.

```{r}
logit2prob <- function(x){
  exp(x) / (1 + exp(x))
}
prob2odds <- function(x){
  x / (1 - x)
}
prob2logodds <- function(x){
  log(prob2odds(x))
}

# augment the dataset using the defined functions
vote_mental_pred <- data1 %>%
  add_predictions(fit_logistic_sim) %>%
  # predicted values are in the log-odds form - convert to probabilities
  mutate(prob = logit2prob(pred)) %>%
  mutate(odds = prob2odds(prob)) %>%
  mutate(logodds = prob2logodds(prob))
```

```{r}
ggplot(vote_mental_pred, aes(x = mhealth_sum)) +
  geom_line(aes(y = logodds), color = "blue", size = 1) +
  labs(title = "Log-odds of voter turnout for different mental health states",
       x = "Mental Health",
       y = "Log-odds of Voter Turnout")
```

# BM - 3

The figure represents the relationship between odds and the mental health. For a healthy mental person, the probability of voting is three times the probability of not voting. On the other hand, for a mentally unhealthy person (mental health about 15), the possibility of not voting is about three times the probability of voting.

```{r}
ggplot(vote_mental_pred, aes(x = mhealth_sum)) +
  geom_line(aes(y = odds), color = "red", size = 1) +
  labs(title = "Odds of Voter Turnout for Different Mental Health States",
       x = "Mental Health",
       y = "Odds Voter Turnout")

geom_point(aes(y = vote96))
```

# BM - 4

The change in probability depends on the initial values. Let's see some examples.

```{r}
ggplot(vote_mental_pred, aes(x = mhealth_sum)) +
#  geom_point(aes(y = vote96)) +
  geom_jitter(aes(y=vote96), na.rm=TRUE, height=.03, width=.3, alpha=.3) +
  geom_line(aes(y = prob), color = "black", size = 1) +
  labs(title = "Probability of voter turout for different mental health states",
       subtitle = '0 = no, 1 = yes; missing values are removed',
       x = "Mental health",
       y = "Probability of voter turnout")


exponential_part_2 <- exp(1.13921 + -0.14348 * 2)
exponential_part_1 <- exp(1.13921 + -0.14348 * 1)
change2_1 <- ((exponential_part_2)/(1+exponential_part_2) - (exponential_part_1)/(1+exponential_part_1))

exponential_part_6 <- exp(1.13921 + -0.14348 * 6)
exponential_part_5 <- exp(1.13921 + -0.14348 * 5)
change5_6 <- ((exponential_part_6)/(1+exponential_part_6) - (exponential_part_5)/(1+exponential_part_5))

cat("The difference in probability for an increase in the mental health index from 1 to 2 is", change2_1)
cat("The difference in probability for an increase in the mental health index from 5 to 6 is", change5_6)

```

# BM - 5 

It is not a good model. The performance of this is worst than a naive assumption of everyone vote, in the case of the accuracy rate. The PRE is quite small, just 0.016, and the AUC supports the conclusion that this model is not the best (0.62).

```{r}
accuracy <- data1 %>%
  add_predictions(fit_logistic_sim) %>%
  mutate(pred = logit2prob(pred),
         prob = pred,
         pred = as.numeric(pred > .5))

accuracy_rate <- mean(accuracy$vote96 == accuracy$pred, na.rm = TRUE)
cat("Accuracy rate:", accuracy_rate)

# function to calculate PRE for a logistic regression model
PRE <- function(model){
  # get the actual values for y from the data
  y <- model$y

  # get the predicted values for y from the model
  y.hat <- round(model$fitted.values)

  # calculate the errors for the null model and your model
  E1 <- sum(y != median(y))
  E2 <- sum(y != y.hat)

  # calculate the proportional reduction in error
  PRE <- (E1 - E2) / E1
  return(PRE)
}

cat("Proportional reduction in error:", PRE(fit_logistic_sim))

# auc <- auc(accuracy$vote96, accuracy$prob)
# auc
```

## Multiple Variable Model

# MVM 1

1.1. Probability distribution: Bernoulli,

Pr$(Y_i = y_i | \pi)$ = $\pi_i^{y_i}$ $(1 - \pi_i)^{1-y_i}$

1.2. Linear predictor: 

$\eta_i = \beta_0 + \beta_1 X_{1,i} + \beta_2 X_{2,i} + \beta_3 X_{3,i} + \beta_4 X_{4,i} + \beta_5 X_{5,i} + \beta_6 X_{6,i} + \beta_7 X_{7,i}$

$X_1$ is mhealth, $X_2$ is age, $X_3$ is educ, $X_4$ is black, $X_5$ is female, $X_6$ is married, and $X_7$ is inc10.

1.3. 

Link function: logit, 

$\pi_i$ $=$ $e^{\eta_i}$ / (1 + $e^{\eta_i}$)

# MVM 2

Results

```{r}
fit_logistic_mul <- glm(vote96 ~ ., data=data1, family=binomial())
summary(fit_logistic_mul)
```

# MVM 3

Interpret the results

Only five variables are significant at the standard level of significance (0.05), including the intercept. These results are in log-odds. The significant variables where mental health, age, education, and income. To make them more easy to interpret. I will utilize a base case, in the median and mode of the variables.

Even controlling for all the variables in the set, the relationship between vote probability and mental health is high. The change in probability in the median and mode of the variables, in a change of mental health from one to two is -0.02. 

The accuracy rate is still modest against the simplified case. Then, in this indicator, it is not a very good model. Also, in the proportional reduction in error, the model perform better than the former model. However, a reduction of 15% in error is still modest.

The following graphs depict some features of the model. For instance, black and married persons will have more probability to vote, nonetheless their mental health. Another interesting characteristic is that the difference between male and female is small. Finally, people below 30 years old tend to vote less frequently than people above 30 years old.

```{r}
median(data1$mhealth_sum, na.rm=TRUE)
median(data1$age, na.rm=TRUE)
median(data1$educ, na.rm=TRUE)
mlv(data1$black, method = "mfv", na.rm=TRUE)
mlv(data1$female, method = "mfv", na.rm=TRUE)
mlv(data1$married, method = "mfv", na.rm=TRUE)
median(data1$inc10, na.rm=TRUE)
```


```{r}
exp_part_2 = exp(-4.304103 + (2 * -0.089102) + 42*0.042534 + 13*0.228686 + 0*0.272984 + 1*-0.016969 + 0*0.296915 + 3.4774*0.069614)

exp_part_1 = exp(-4.304103 + (1 * -0.089102) + 42*0.042534 + 13*0.228686 + 0*0.272984 + 1*-0.016969 + 0*0.296915 + 3.4774*0.069614)

prob_p2 <- exp_part_2/(1+exp_part_2)
prob_p1 <- exp_part_1/(1+exp_part_1)

cat("The change in probability for an change of mental health from 1 to 2 is:", prob_p2 - prob_p1)

```


```{r}
accuracy_mul <- data1 %>%
  add_predictions(fit_logistic_mul) %>%
  mutate(pred = logit2prob(pred),
         prob = pred,
         pred = as.numeric(pred > .5))

accuracy_rate_mul <- mean(accuracy_mul$vote96 == accuracy_mul$pred, na.rm = TRUE)
cat("Accuracy rate:", accuracy_rate_mul);
cat("Proportional reduction in error:", PRE(fit_logistic_mul));
cat("Area Under the Curve:", auc(accuracy_mul$vote96, accuracy_mul$prob))
```


```{r}
vote_mental_pred2 <- data1 %>%
  data_grid(mhealth_sum, black, .model=fit_logistic_mul) %>%
  add_predictions(fit_logistic_mul) %>%
  # predicted values are in the log-odds form - convert to probabilities
  mutate(prob = logit2prob(pred))

ggplot(vote_mental_pred2, aes(x = mhealth_sum, y = prob, color = ifelse(black > 0.5 , "Black", "Non Black"))) +
  geom_smooth() +
  labs(title = "Probability of voter turnout for different mental health states",
       subtitle = '0 = no, 1 = yes; missing values are removed',
       x = "Mental Health",
       y = "Probability of Voter Turnout") +
  guides(color = guide_legend(''))
```


```{r}
vote_mental_pred2 <- data1 %>%
  data_grid(mhealth_sum, female, .model=fit_logistic_mul) %>%
  add_predictions(fit_logistic_mul) %>%
  # predicted values are in the log-odds form - convert to probabilities
  mutate(prob = logit2prob(pred))

ggplot(vote_mental_pred2, aes(x = mhealth_sum, y = prob, color = ifelse(female > 0.5 , "Female", "Male"))) +
  geom_smooth() +
  labs(title = "Probability of voter turnout for different mental health states",
       subtitle = '0 = no, 1 = yes; missing values are removed',
       x = "Mental Health",
       y = "Probability of Voter Turnout") +
  guides(color = guide_legend(''))
```

```{r}
vote_mental_pred2 <- data1 %>%
  data_grid(mhealth_sum, married, .model=fit_logistic_mul) %>%
  add_predictions(fit_logistic_mul) %>%
  # predicted values are in the log-odds form - convert to probabilities
  mutate(prob = logit2prob(pred))

ggplot(vote_mental_pred2, aes(x = mhealth_sum, y = prob, color = ifelse(married > 0.5 , "Married", "Single"))) +
  geom_smooth() +
  labs(title = "Probability of voter turnout for different mental health states",
       subtitle = '0 = no, 1 = yes; missing values are removed',
       x = "Mental Health",
       y = "Probability of Voter Turnout") +
  guides(color = guide_legend(''))
```

```{r}
vote_mental_pred2 <- data1 %>%
  data_grid(mhealth_sum, age, .model=fit_logistic_mul) %>%
  add_predictions(fit_logistic_mul) %>%
  # predicted values are in the log-odds form - convert to probabilities
  mutate(prob = logit2prob(pred))

ggplot(vote_mental_pred2, aes(x = mhealth_sum, y = prob, color = ifelse(age < 31 , "Young", "Other"))) +
  geom_smooth() +
  labs(title = "Probability of voter turnout for different mental health states",
       subtitle = '0 = no, 1 = yes; missing values are removed',
       x = "Mental Health",
       y = "Probability of Voter Turnout") +
  guides(color = guide_legend(''))
```


## TV Model

# TV - 1
1. Components

1.1. Probability: the Poisson distribution, 

Pr$(Y_i = y_i | \mu)$ = $\mu^{y_i}$ $e^{-\mu}$ / $y_i!$

1.2. Linear predictor: 

$\eta_i = \beta_0 + \beta_1 X_{1,i} + \beta_2 X_{2,i} + \beta_3 X_{3,i} + \beta_4 X_{4,i} + \beta_5 X_{5,i} + \beta_6 X_{6,i} + \beta_7 X_{7,i}$,
where $X_1$ is age, $X_2$ is educ, $X_3$ is female, $X_4$ is hours per day to relax, $X_5$ is black, $X_6$ is voted in the 2004 election, and $X_7$ is x-rated movie.

1.3. Link function: the log, 

$\mu_i$ $=$ ln($\eta_i$)

# TV - 2

```{r}
fit_poisson_mul2 <- glm(tvhours ~ age +educ +female +hrsrelax +black +voted04 +xmovie, data=data2, family=poisson())
summary(fit_poisson_mul2)
```

# TV - 3

```{r}
# Interpretation
educ_int = exp(-0.0274218)
hrealx_int = exp(0.0484390)
black_int = exp(0.4294380)

cat("The change in watching tv by year of education is", educ_int)
cat("The change in watching tv by hour of relax is", hrealx_int)
cat("The change in watching tv for black against others is", black_int)
```

The model predicts some intuitive results, such as people with more free time will watch more TV, ot that black people (associated with low income) will watch more TV. However, the coefficient for education appears to be wrong because more education is correlated with more time spending watching TV. Then, there are three conclusions about this, more research is needed to understand this relationship, the data is bias on some way, or education is correlated with another ommited variable that can explain the results. 

The regression using the quasi-Poisson method shows that the dispersion parameter is close to 1. Then, underdispersion does not appear to be a significant problem.
```{r}
fit_quasi <- glm(formula = tvhours ~ age +educ +female +hrsrelax +black +voted04 +xmovie, data=data2, family="quasipoisson")
summary(fit_quasi)
```

