---
title: "Problem set #8: Trees and SVMs"
author: "Sushmita V Gopalan"
output:
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
library(forcats)
library(broom)
library(modelr)
library(tree)
library(randomForest)
library(stringr)
library(ISLR)
library(gridExtra)
library(grid)
library(titanic)
#library(rcfss)
library(pROC)
library(gbm)
library(caret)
library(e1071)
library(ggdendro)
options(digits = 3)

theme_set(theme_minimal())

# read in data
biden <- read_csv("biden.csv")
```

## Question 1 - Joe Biden

** Split Dataset **

[1] Split the data into a training set (70%) and a validation set (30%). Be sure to set your seed prior to this part of your code to guarantee reproducibility of results.

```{r}
set.seed(1234) # For reproducibility
biden.split <- resample_partition(biden, c(test = .3, train = .7))
```

**********************************************************

** Basic Decision Tree **

[2] Fit a decision tree to the training data, with biden as the response variable and the other variables as predictors. Plot the tree and interpret the results. What is the test MSE?

```{r}
# estimate model
biden_tree <- tree(biden ~ ., data = biden.split$train)

# plot tree
tree_data <- dendro_data(biden_tree)
ptree <- ggplot(segment(tree_data)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), 
               alpha = 0.5) +
  geom_text(data = label(tree_data), 
            aes(x = x, y = y, label = label_full), vjust = -0.5, size = 3) +
  geom_text(data = leaf_label(tree_data), 
            aes(x = x, y = y, label = label), vjust = 0.5, size = 3) +
  theme_dendro()+
  labs(title = 'Decision Tree for Biden Scores',
       subtitle = 'All predictors, Default Controls')

# function to get MSE
mse <- function(model, data) {
  x <- modelr:::residuals(model, data)
  mean(x ^ 2, na.rm = TRUE)
}

mse_biden_1 = mse(biden_tree,biden.split$test)
leaf_vals <- leaf_label(tree_data)$yval
ptree


```

Having only default tree controls in the model makes it straightforward to interpret. Only 'dem' and 'rep' were used to predict Biden Score. 

If the value of `dem` > 0.5, in our case, `dem` = 1, i.e. a person is a democrat, the model predicts a Biden score of `r leaf_vals[3]` 

If the value of `dem` is < 0.5, i.e. the person is either a Republican or neither, we go down the left side of the decision tree. Here we see that if `rep` > 0.5, we go down the right side of that sub-tree to find that the model predicts a Biden value of `r leaf_vals[2]`
We go down the left of side of the sub-tree to find that, for individuals who are neither Republican nor Democrat, our model predicts a Biden score of `r leaf_vals[1]`

The value of the MSE is `r mse_biden_1`

**********************************************************

** Decision Tree with Pruning**

[3] Now fit another tree to the training data with the following control options:

tree(control = tree.control(nobs = # number of rows in the training set,mindev = 0))

Use cross-validation to determine the optimal level of tree complexity, plot the optimal tree, and interpret the results. Does pruning the tree improve the test MSE?

```{r}
set.seed(1234) # For reproducibility

biden_tree_2 <- tree(biden ~ ., data = biden.split$train,
     control = tree.control(nobs = nrow(biden.split$train),
                            mindev = 0))
mod <- biden_tree_2

mse_biden_2 <- mse(biden_tree_2, biden.split$test)

num_nodes <- 2:25
pruned_trees <- map(num_nodes, prune.tree, tree = biden_tree_2, k = NULL)
test_mses <- map_dbl(pruned_trees, mse, data = biden.split$test)

tree.opt <- pruned_trees[[which.min(test_mses)]]
opt_test_mse <- mse(tree.opt, biden.split$test)

tibble(num_nodes = num_nodes, test_mse = test_mses) %>%
  ggplot(aes(x = num_nodes, y = test_mse)) +
  geom_line() + 
  labs(title = 'Test MSE for Different Numbers of Terminal Nodes',
       subtitle = '(Test MSE calculated on Test data defined in Step 1)',
       x = 'Terminal Nodes in Tree',
       y = 'Test MSE')


biden_pruned <- prune.tree(biden_tree_2, best=11)
mse_pruned = mse(biden_pruned,biden.split$test)

```
 
Using cross validation, we find from this graph that the MSE is lowest for a tree with 11 nodes. Pruning to 11 nodes reduces the MSE from `r mse_biden_2` to `r mse_pruned`. Here's what a plot of the tree with 11 nodes looks like -
 
```{r}
plot(biden_pruned, col='purple', lwd=2.5)
title("Best 11 Regression Tree for Biden Scores\n")
text(biden_pruned, col='red')
```

The model suggests that for democrats, age is the most important predictor of Biden scores, followed by education. Gender does not appear to influence Biden scores.

For Republicans too, gender does not appear to be important in predicting Biden score. Age is again, the most important predictor. In the 44-47 age group, education also influences Biden scores.

For the unaffiliated, gender is an important predictor. For women, education and age influence Biden scores, but they do not for men.

**********************************************************

** Bagging **

[4] Use the bagging approach to analyze this data. What test MSE do you obtain? Obtain variable importance measures and interpret the results.


```{r}
# prep data for bagging
# defining new datasets for train and test because the 'select' method
# can't be applied to an object of class 'resample'

df = read.csv('biden.csv')
df$Party[df$dem == 1] = 'Democrat'
df$Party[df$dem == 0 & df$rep == 0] = 'No Affiliation'
df$Party[df$rep == 1] = 'Republican'

set.seed(1234)

biden_split7030 = resample_partition(df, c(test = 0.3, train = 0.7))
biden_train70 = biden_split7030$train %>%
                tbl_df()
biden_test30 = biden_split7030$test %>%
               tbl_df()

biden_bag_data_train = biden_train70 %>%
                       select(-Party) %>%
                       mutate_each(funs(as.factor(.)), dem, rep) %>%
                       na.omit

biden_bag_data_test = biden_test30 %>%
                      select(-Party) %>%
                      mutate_each(funs(as.factor(.)), dem, rep) %>%
                      na.omit

# estimate model
(bag_biden <- randomForest(biden ~ ., data = biden_bag_data_train, mtry = 5, ntree = 500, importance=TRUE))
# find MSE
mse_bag_biden = mse(bag_biden, biden_bag_data_test)
``` 


The MSE for the model with bagging is `r mse_bag_biden` , which is much higher than we had for the pruned tree with `r mse_pruned` 
The % variation explained is also very low, at 8.91%.



```{r}
set.seed(1234)

bag_biden_importance = as.data.frame(importance(bag_biden))

ggplot(bag_biden_importance, mapping=aes(x=rownames(bag_biden_importance), y=IncNodePurity)) +
       geom_bar(stat="identity", aes(fill=IncNodePurity)) +
       labs(title = "Average Increased Node Purity Across 500 Regression Trees",
       subtitle = "Predicted Warmth Toward Joe Biden (2008)",
       x = "Variable",
       y = "Mean Increased Node Purity") + 
       theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
             panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1), legend.position = 'none') 
```

As per the above graph, we see that `age` and `dem` are the variables which produce the highest average decreases in node impurity across 500 bagged regression trees. The bagging model uses bootstrapping to create 500 different training datasets, while the pruned tree uses only one set of training data. The bagging model looks at the mean variance across all the bootstrapped trees and estimates that 'age' and `dem` are the most indicative variables in the model, while 'female' is the least.

**********************************************************
** Random Forest **

[5] Use the random forest approach to analyze this data. What test MSE do you obtain? Obtain variable importance measures and interpret the results. Describe the effect of m, the number of variables considered at each split, on the error rate obtained.

```{r}
# estimate RF model
set.seed(1234)

(biden_rf <- randomForest(biden ~ ., data = biden_bag_data_train,mtry =2,ntree = 500))
mse_rf = mse(biden_rf, biden_bag_data_test)

# try to plot a graph that shows MSE is minimum for mtry = 2
``` 

```{r}
set.seed(1234)

rf_biden_importance = as.data.frame(importance(biden_rf))

ggplot(rf_biden_importance, mapping=aes(x=rownames(rf_biden_importance), y=IncNodePurity)) +
       geom_bar(stat="identity", aes(fill=IncNodePurity)) +
       labs(title = "Average Increased Node Purity Across 500 Regression Trees",
       subtitle = "Predicted Warmth Toward Joe Biden (2008)",
       x = "Variable",
       y = "Mean Increased Node Purity") + 
       theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
             panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1), legend.position = 'none') 
```

The MSE for this model is `r mse_rf` which we see is much less than `r mse_bag_biden`.
The random forest model with mtry = 2 suggests that 'dem' and 'rep' are the most important variables in the model, i.e. they yield the highest decreases in node impurity. 

Random forests offer an improvement over bagged trees by 'decorrelating' the trees. We still build a number of decision trees on bootstrapped training samples. In a random forest model, each time a split in a tree is considered, a random sample of m predictors is chosen as split candidates from the full set of p predictors. We restrict ourselves to a random subset m of all predictors in order to avoid the possibility of all trees looking extremely similar like in a bagged model, where all the trees would end up using the same strongest predictor in the top split. In such a situation, averaging across highly correlated predictions would not lead to a substantial decrease in variance. In the random forest model, we overcome this issue by restricting each split to only a subset of predictors - basically, about (p-m)/p of the splits wouldn't even consider the strongest predictor, allowing the influence of the other predictors to become discernible. Conventionally, using m approximately equal to the square root of p leads to a reduction in both test error and OOB error of bagging.

**********************************************************
** Boosting **

[6] Use the boosting approach to analyze the data. What test MSE do you obtain? How does the value of the shrinkage parameter λ influence the test MSE?

We first run the boosting model using depths of 1,2 and 4 respoectively, to find the optimal number of iterations for lowest MSE.
```{r,echo=FALSE,message=FALSE,warning=FALSE}
# plot all models
set.seed(1234)
biden_models <- list("boosting_depth1" = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 10000, interaction.depth = 1),
                       "boosting_depth2" = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 10000, interaction.depth = 2),
                       "boosting_depth4" = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 10000, interaction.depth = 4))
```

```{r}
set.seed(1234)
data_frame(depth = c(1, 2, 4),
           model = biden_models[c("boosting_depth1", "boosting_depth2", "boosting_depth4")],
           optimal = map_dbl(model, gbm.perf, plot.it = FALSE)) %>%
  select(-model) %>%
  knitr::kable(caption = "Optimal number of boosting iterations",
               col.names = c("Depth", "Optimal number of iterations"))
```

```{r} 
# default shrinkage
set.seed(1234)

biden_boost_1 = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 3302, interaction.depth = 1)

biden_boost_2 = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 2700, interaction.depth = 2)

biden_boost_4 = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 2094, interaction.depth = 4)
```
Using these optimal values, we run the boost model again to see what depth at its ideal number of trees gives us lowest MSE.

```{r,echo=FALSE}
# get MSE
set.seed(1234)
## work around bug in gbm 2.1.1
# obtained from http://www.samuelbosch.com/2015/09/workaround-ntrees-is-missing-in-r.html

predict.gbm <- function (object, newdata, n.trees, type = "link", single.tree = FALSE, ...) {
  if (missing(n.trees)) {
    if (object$train.fraction < 1) {
      n.trees <- gbm.perf(object, method = "test", plot.it = FALSE)
    }
    else if (!is.null(object$cv.error)) {
      n.trees <- gbm.perf(object, method = "cv", plot.it = FALSE)
    }
    else {
      n.trees <- length(object$train.error)
    }
    cat(paste("Using", n.trees, "trees...\n"))
    gbm::predict.gbm(object, newdata, n.trees, type, single.tree, ...)
  }
}
mse_1 = mse(biden_boost_1,biden_bag_data_test)
mse_2 = mse(biden_boost_2,biden_bag_data_test)
mse_4 = mse(biden_boost_4,biden_bag_data_test)
``` 
MSE_1 = `mse_1`
MSE_2 = `mse_2`
MSE_4 = `mse_4`
We see that using a depth of 1 with 3302 trees gives us an MSE value that is lowest, although only marginal less than (depth = 2, trees = 2700) and (depth = 4, trees = 2094).

We also see that these MSE values are much better than those obtained by our bagging and random forest models. 

Let's see what happens when we change the default shrinkage of lambda = 0.001 to lambda = 0.0005  

```{r} 
# default shrinkage
set.seed(1234)

biden_boost_1_2 = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 3302, interaction.depth = 1,shrinkage=0.0005)

biden_boost_2_2 = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 2700, interaction.depth = 2,shrinkage=0.0005)

biden_boost_4_2 = gbm(as.numeric(biden) - 1 ~ .,
                                               data = biden_bag_data_train,
                                               n.trees = 2094, interaction.depth = 4,shrinkage=0.0005)
```

Using these optimal values, we run the boost model again to see what depth at its ideal number of trees gives us lowest MSE.
```{r,echo=FALSE}
# get MSE

mse_1_2 = mse(biden_boost_1_2,biden_bag_data_test)
mse_2_2 = mse(biden_boost_2_2,biden_bag_data_test)
mse_4_2 = mse(biden_boost_4_2,biden_bag_data_test)
``` 
MSE_1 = `mse_1_2`
MSE_2 = `mse_2_2`
MSE_4 = `mse_4_2`

We notice that MSE values have increased. This makes sense because given the same number of trees, a smaller value of lambda increases MSE. Lambda, essentially, is a shrinkage parameter which controls the rate at which boosting learns. So, a very small value of lambda will require a very large number of trees in order to achieve good performance. 

## Question 2 - Voting Patterns

```{r mentalprep,echo=FALSE, }
mh <- read_csv("mental_health.csv") %>%
  mutate_each(funs(as.factor(.)), vote96, black, female, married) %>%
  na.omit

set.seed(5678)
mh_split <- resample_partition(mh, p = c("test" = .3, "train" = .7))
```

[1]Use cross-validation techniques and standard measures of model fit (e.g. test error rate, PRE, ROC curves/AUC) to compare and evaluate at least five tree-based models of voter turnout. Select the best model and interpret the results using whatever methods you see fit (graphs, tables, model fit statistics, predictions for hypothetical observations, etc.)```{r model1}



```{r}

mh_tree <- tree(vote96 ~ educ, data = as_tibble(mh_split$train))
mh_tree

fitted <- predict(mh_tree, as_tibble(mh_split$test), type = "class")
tree_err <- mean(as_tibble(mh_split$test)$vote96 != fitted)
tree_err

roc_tree1 <- roc(as.numeric(as_tibble(mh_split$test)$vote96), as.numeric(fitted))
plot(roc_tree1)

auc(roc_tree1)
```

```{r model2}
mh_tree <- tree(vote96 ~ educ + mhealth_sum, data = as_tibble(mh_split$train))
mh_tree

plot(mh_tree)
text(mh_tree, pretty = 0)

fitted <- predict(mh_tree, as_tibble(mh_split$test), type = "class")
tree_err <- mean(as_tibble(mh_split$test)$vote96 != fitted)
tree_err

roc_tree2 <- roc(as.numeric(as_tibble(mh_split$test)$vote96), as.numeric(fitted))
plot(roc_tree2)

auc(roc_tree2)
```

```{r model3}
mh_tree <- tree(vote96 ~ educ + mhealth_sum + age, data = as_tibble(mh_split$train))
mh_tree

plot(mh_tree)
text(mh_tree, pretty = 0)

fitted <- predict(mh_tree, as_tibble(mh_split$test), type = "class")
tree_err <- mean(as_tibble(mh_split$test)$vote96 != fitted)
tree_err

roc_tree3 <- roc(as.numeric(as_tibble(mh_split$test)$vote96), as.numeric(fitted))
plot(roc_tree3)

auc(roc_tree3)
```

```{r model4}
mh_tree <- tree(vote96 ~ educ + mhealth_sum + age + inc10, data = as_tibble(mh_split$train))
mh_tree

plot(mh_tree)
text(mh_tree, pretty = 0)

fitted <- predict(mh_tree, as_tibble(mh_split$test), type = "class")
tree_err <- mean(as_tibble(mh_split$test)$vote96 != fitted)
tree_err

roc_tree4 <- roc(as.numeric(as_tibble(mh_split$test)$vote96), as.numeric(fitted))
plot(roc_tree4)

auc(roc_tree4)
```

```{r model5}
mh_tree <- tree(vote96 ~ ., data = as_tibble(mh_split$train))
mh_tree

plot(mh_tree)
text(mh_tree, pretty = 0)

fitted <- predict(mh_tree, as_tibble(mh_split$test), type = "class")
tree_err <- mean(as_tibble(mh_split$test)$vote96 != fitted)
tree_err

roc_tree5 <- roc(as.numeric(as_tibble(mh_split$test)$vote96), as.numeric(fitted))
plot(roc_tree5)

auc(roc_tree5)
```

```{r compare_trees}
plot(roc_tree1, print.auc = TRUE, col = "blue", print.auc.x = .2)
plot(roc_tree2, print.auc = TRUE, col = "red", print.auc.x = .2, print.auc.y = .4, add = TRUE)
plot(roc_tree3, print.auc = TRUE, col = "orange", print.auc.x = .2, print.auc.y = .3, add = TRUE)
plot(roc_tree4, print.auc = TRUE, col = "green", print.auc.x = .2, print.auc.y = .2, add = TRUE)
plot(roc_tree5, print.auc = TRUE, col = "purple", print.auc.x = .2, print.auc.y = .1, add = TRUE)
```

The models I picked varied in terms of the predictor variables included.
Model 1 contains only the predictor variable `educ`.
Model 2 contains `educ` and `mhealth_sum`.
Model 3 contains `educ`, `mhealth_sum`, and `age`.
Model 4 contains `educ`, `mhealth_sum`, `age` and `inc10`
Model 5 contains all possible predictor values.

The areas under the curve and the test tree errors are exactly the same for models 3, 4, and 5. From the consolidated graph, we see that the AUC for these three models is the highest at 0.686. 

Adding the variables `black`, `female`, `married`, and `inc10` neither increases AUC nor decreases the error rate, leading me to conclude that model 3 is the best model. It has highest AUC and lowest error, without adding variables that increase computational cost without offering more predictive insight 

Let's examine model 3 in closer detail - 

```{r besttree}
mh_tree <- tree(vote96 ~ educ + mhealth_sum + age, data = as_tibble(mh_split$train))
mh_tree

plot(mh_tree)
text(mh_tree, pretty = 0)
```

In model 3, age is the most important predictor, followed by education and mental health indeex. The depiction of the decision tree above helps to flesh out the predictions - for instance, consider an individual over the age of 45. With `educ` > 12.5, the individual is predicted to have voted, irrespective of their `mhhealth` index. However, with `educ` < 12.5, we see that `mhealth` becomes important - only those with `mhealth` < 4.5 are predicted to have voted, while others are not. 

Traveling down the left side of the tree, we observe that for individuals with `age` < 45, `educ` is the next most important predictor. Those with `educ` < 13.5 are predicted not to have voted at all, irrespetive of their mental health status! For those with `educ` > 13.5, `mhealth` becomes important, with those with `mhealth` scores > 3.5 predicted not to vote and those with lower `mhealth` scores are predicted to vote. 

** Part 2: SVMs ** 
[2] Use cross-validation techniques and standard measures of model fit (e.g. test error rate, PRE, ROC curves/AUC) to compare and evaluate at least five SVM models of voter turnout. Select the best model and interpret the results using whatever methods you see fit (graphs, tables, model fit statistics, predictions for hypothetical observations, etc.)

```{r}
set.seed(1234)
mh_data = read.csv("mental_health.csv") %>%
  mutate_each(funs(as.factor(.)), vote96, black, female, married) %>%
  na.omit

mh_split <- resample_partition(mh_data, p = c("test" = .3, "train" = .7))

```

** Model 1: Linear Kernel with education, age, mhealthsum **

```{r}
mh_lin_tune <- tune(svm, vote96 ~ educ + age + mhealth_sum, data = as_tibble(mh_split$train),
                    kernel = "linear",
                    range = list(cost = c(.001, .01, .1, 1, 5, 10, 100)))

mh_lin <- mh_lin_tune$best.model
summary(mh_lin)

```


```{r}
fitted <- predict(mh_lin, as_tibble(mh_split$test), decision.values = TRUE) %>%
  attributes


roc_line <- roc(as_tibble(mh_split$test)$vote96, fitted$decision.values)

auc(roc_line)
plot(roc_line, main = "ROC of Voter Turnout - Linear Kernel, Partial Model")

```
The area under the curve is 0.737.

** Model 2: Linear Kernel with all variables **
```{r}
mh_lin_all <- tune(svm, vote96 ~ ., data = as_tibble(mh_split$train),
                    kernel = "linear",
                    range = list(cost = c(.001, .01, .1, 1, 5, 10, 100)))

mh_lall <- mh_lin_all$best.model
summary(mh_lall)

```

```{r}
fitted <- predict(mh_lall, as_tibble(mh_split$test), decision.values = TRUE) %>%
  attributes


roc_line_all <- roc(as_tibble(mh_split$test)$vote96, fitted$decision.values)

auc(roc_line_all)
plot(roc_line_all, main = "ROC of Voter Turnout- Linear Kernel, Total Model")

```
Area under the curve is 0.746.

** Model 3: Polynomial Kernel with education, age, mhealth **
```{r}
mh_poly_tune <- tune(svm, vote96 ~ age + educ + mhealth_sum, data = as_tibble(mh_split$train),
                    kernel = "polynomial",
                    range = list(cost = c(.001, .01, .1, 1, 5, 10, 100)))

mh_poly <- mh_poly_tune$best.model
summary(mh_poly)

```

```{r}
fitted <- predict(mh_poly, as_tibble(mh_split$test), decision.values = TRUE) %>%
  attributes

roc_poly <- roc(as_tibble(mh_split$test)$vote96, fitted$decision.values)
plot(roc_poly, main = "ROC of Voter Turnout - Polynomial Kernel, Partial Model")

```
The area under the curve is 0.743. 

** Model 4: Polynomial Model A **

```{r}
mh_poly_all <- tune(svm, vote96 ~ ., data = as_tibble(mh_split$train),
                    kernel = "polynomial",
                    range = list(cost = c(.001, .01, .1, 1, 5, 10, 100)))

mh_poly <- mh_poly_all$best.model
summary(mh_poly)

```

```{r}
fitted <- predict(mh_poly, as_tibble(mh_split$test), decision.values = TRUE) %>%
  attributes

roc_poly_all <- roc(as_tibble(mh_split$test)$vote96, fitted$decision.values)
plot(roc_poly_all, main = "ROC of Voter Turnout - Polynomial Kernel, Total Model")

```
Area under the curve: 0.741.

** Model 5: Radial Kernel **

```{r}
mh_rad_tune <- tune(svm, vote96 ~ ., data = as_tibble(mh_split$train),
                    kernel = "radial",
                    range = list(cost = c(.001, .01, .1, 1, 5, 10, 100)))

mh_rad <- mh_rad_tune$best.model
summary(mh_rad)

```
```{r}
fitted <- predict(mh_rad, as_tibble(mh_split$test), decision.values = TRUE) %>%
  attributes

roc_rad <- roc(as_tibble(mh_split$test)$vote96, fitted$decision.values)
plot(roc_rad, main= "ROC of Voter Turnout - Radial Kernel, Total Model")

```
Area under the curve is 0.737.

```{r}
plot(roc_line, print.auc = TRUE, col = "blue", print.auc.x = .2)
plot(roc_line_all, print.auc = TRUE, col = "red", print.auc.x = .2, print.auc.y = .4, add = TRUE)
plot(roc_poly, print.auc = TRUE, col = "orange", print.auc.x = .2, print.auc.y = .3, add = TRUE)
plot(roc_poly_all, print.auc = TRUE, col = "green", print.auc.x = .2, print.auc.y = .2, add = TRUE)
plot(roc_rad, print.auc = TRUE, col = "purple", print.auc.x = .2, print.auc.y = .1, add = TRUE)

```

The best model is model 1- the linear kernel with only education, age, and mental health. With highest area under the curve so it offers the most improvement over the useless classifier. This model has a cost of 1, so the margins are narrow around the linear hyperplane. The graph below shows that the error drops to below 0.3 and stays there, regardless of the increases in error. 

```{r}
plot(mh_lin_tune)

```

```{r}
svm(vote96 ~ age + educ + mhealth_sum, data = as_tibble(mh_split$test), kernel = "linear", scale = FALSE, cost = 1) %>%
  plot(as_tibble(mh_split$test), age ~ educ)

```

```{r}
svm(vote96 ~ educ + mhealth_sum + age, data = as_tibble(mh_split$test), kernel = "linear", scale = FALSE, cost = 1) %>%
  plot(as_tibble(mh_split$test), educ ~ mhealth_sum)

```

```{r}
svm(vote96 ~ mhealth_sum + age + educ, data = as_tibble(mh_split$test), kernel = "linear", scale = FALSE, cost = 1) %>%
  plot(as_tibble(mh_split$test), mhealth_sum ~ age)

```



## Question 3 - OJ Simpson

```{r,echo=FALSE, include=FALSE}
# read in data 
simpson <- read_csv("simpson.csv")
simpson$male <- with(simpson, ifelse(female == 1, 0, 1))
simpson$white <- with(simpson, ifelse(black == 0 & hispanic == 0,1,0))

```
** Race and OJ's Guilt **

[1] What is the relationship between race and belief of OJ Simpson's guilt? Develop a robust statistical learning model and use this model to explain the impact of an individual's race on their beliefs about OJ Simpson's guilt.

A natural starting point to the model the relationship between race and belief of OJ's guilt would be to use a logistic regression. This is because 'guilt' is a binary variable - taking 0 when the respondent believes OJ is innocent and 1 when they believe he's guilty.

When I tried including 'educ' and 'income' in my model, some dummy variables were found to be weakly significant. Educ being 'not high school grad' had a negative impact on belief in OJ's guilt, having income 'over $75000' had a positive impact on belief in OJ's guily. However, I did not include those variables because the model's overall accuracy did not benefit from their inclusion. 

Including `hispanic` in the model too, resulted in discovering that being `hispanic` does not have a statistically significant impact on belief in OJ's guilt. This points to the idea that within the idea of race, specifically, being black, impacts belief in OJ's guilt.

I also created a dummy variable indicating 'male' and one indicating 'neither black nor hispanic' - to see if there was an interaction effect between between race and gender - it is conceivable that black men are more likely to believe in OJ's innocence (same race-gender category as OJ) or white (roughly approximated by 'neither black nor hispanic') women (same race-gender cateogry as OJ's alleged victim) are less likely to - when I ran the regression however, I was suprised to find that the interaction term was not significant. 

Other tested models:
```{r}
black_educ_income <- glm(guilt ~black+educ+income,data = simpson, family = binomial)
all_vars <- glm(guilt ~black+educ+income+age+rep+dem,data = simpson, family = binomial)
black_male <- glm(guilt ~black*male,data = simpson, family = binomial)
white_female <- glm(guilt ~white*female,data = simpson, family = binomial)
black_age <- glm(guilt~black+age,data=simpson,family=binomial)
hispanic_black <- glm(guilt~black+hispanic,data=simpson,family=binomial)
summary(white_female)
```

Here's what a logistic regression of 'guilt' on 'black' looks like - 

```{r}
black_only <- glm(guilt ~black,data = simpson, family = binomial)
summary(black_only)


logit2prob <- function(x){
  exp(x) / (1 + exp(x))
}
accuracy <- simpson %>%
  add_predictions(black_only) %>%
  mutate(pred = logit2prob(pred),
         pred = as.numeric(pred > .5))

model_accuracy = mean(accuracy$guilt == accuracy$pred, na.rm = TRUE)

PRE <- function(model){
  # get the actual values for y from the data
  y <- model$y

  # get the predicted values for y from the model
   y.hat <- round(model$fitted.values)

  # calculate the errors for the null model and your model
  E1 <- sum(y != median(y))
  E2 <- sum(y != y.hat)

  # calculate the proportional reduction in error
  PRE <- (E1 - E2) / E1
  return(PRE)
}

pre <- PRE(black_only)
auc_x <- auc(accuracy$guilt, accuracy$pred)

model_accuracy
pre
auc_x
```

We see here that the relationship between the variable 'black' and belief in OJ's guilt is highly statistically significant, with a p-value to the order of 10^-16.
It is also substantively significant, in that being black reduces the log-odds of an individuals belief in OJ's guilt by -3.1022, i.e., lowers the likelihood of believing in OJ's guilt by ~ 14%.

Model Accuracy : `r model_accuracy`
Proportional Reduction in Error : `r pre`
Area under Curve : `r auc_x`

The model's accuracy is 81.6%, which is good. The proportional error reduction is 41%, which is substantial. The AUC shows a 0.23 increase over the useless classifier.   

** Predictive Model for OJ's Guilt **

Decision trees are great intuitive concepts for making predictions. Theywork by splitting the observations into a number of regions, and predictions are made based on the mean or mode of the training observations in that region.


```{r}
set.seed(1234) # For reproducibility
oj = read.csv('simpson.csv')
oj = oj[(!is.na(oj$guilt)), ]
oj$Opinion = factor(oj$guilt, levels = c(0,1), labels = c("Innocent", "Guilty"))
```

```{r}
oj_split7030 = resample_partition(oj, c(test = 0.3, train = 0.7))
oj_train70 = oj_split7030$train %>%
                tbl_df()
oj_test30 = oj_split7030$test %>%
               tbl_df()

oj_data_train = oj_train70 %>%
                select(-guilt) %>%
                mutate_each(funs(as.factor(.)), dem, rep) %>%
                na.omit

oj_data_test = oj_test30 %>%
               select(-guilt) %>%
               mutate_each(funs(as.factor(.)), dem, rep) %>%
               na.omit

# estimate model
oj_tree <- tree(Opinion ~ ., data = oj_data_train)

# plot tree
tree_data <- dendro_data(oj_tree)

ptree <- ggplot(segment(tree_data)) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), 
               alpha = 0.5) +
  geom_text(data = label(tree_data), 
            aes(x = x, y = y, label = label_full), vjust = -0.5, size = 3) +
  geom_text(data = leaf_label(tree_data), 
            aes(x = x, y = y, label = label), vjust = 0.5, size = 3) +
  theme_dendro()+
  labs(title = "Decision Tree for OJ's Guilt",
       subtitle = 'All predictors, Default Controls')
ptree
```
A rudimentary decision tree confirms what we learned in part 1 - that race is the most important predictor of an individual's perception of OJ Simpson's guilt, with age being a distant second. We will now use a random forest to explore this further. I choose random forests over pruned trees bagging because it averages across a set of uncorrelated trees and reduces variance.
```{r}
# Let's build a random forest
#biden.split <- resample_partition(guilt, c(test = .3, train = .7))
# function to get MSE
mse <- function(model, data) {
  x <- modelr:::residuals(model, data)
  mean(x ^ 2, na.rm = TRUE)
}

(rf_oj = randomForest(Opinion ~ ., data = oj_data_train, mtry = 3,ntree = 500))
```
We notice an error rate of 19.7%, which is a success rate of ~ 80% and that's reasonably good.

```{r}
rf_oj_importance = as.data.frame(importance(rf_oj))

ggplot(rf_oj_importance, mapping=aes(x=rownames(rf_oj_importance), y=MeanDecreaseGini)) +
       geom_bar(stat="identity", aes(fill=MeanDecreaseGini)) + 
       labs(title = "Mean Decrease in Gini Index Across 500 Random Forest Regression Trees",
       subtitle = "Predicted Opinion of Simpson Guilt",
       x = "Variable",
       y = "Mean Decrease in Gini Index") + 
       theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
             panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1), legend.position = 'none') 
```
This graph further confirms that race is the most important predictor of 'Opinion' - it yields the highest decrease in Gini Index. Its importance is followed by 'age' - just as we noticed in our logistic regression model above. The difference in the height of the bars for `black` and `hispanic` further confirms that it is specifically, `black` that influences perception of OJ's guilt - being `hispanic` has no discernible effect. 