---
title: "MACS 30100 PS8"
author: "Erin M. Ochoa"


date: "2017 March 6"
output:
  github_document:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)

library(tidyverse)
library(forcats)
library(broom)
library(modelr)
library(tree)
library(randomForest)
library(stringr)
library(ISLR)
library(gridExtra)
library(grid)
#library(rcfss) #Not available for 3.3.2
library(pROC)
library(gbm)
library(ggdendro) #boltoff's fork


options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())
```

```{r functions}

mse = function(model, data) {
  x = modelr:::residuals(model, data)
  mean(x ^ 2, na.rm = TRUE)
}


# There seems to be a bug in the gbm function.
# Work-around method found here: http://www.samuelbosch.com/2015/09/workaround-ntrees-is-missing-in-r.html

predict.gbm = function (object, newdata, n.trees, type = "link", single.tree = FALSE, ...) {
  if (missing(n.trees)) {
    if (object$train.fraction < 1) {
      n.trees = gbm.perf(object, method = "test", plot.it = FALSE)
    }
    else if (!is.null(object$cv.error)) {
      n.trees = gbm.perf(object, method = "cv", plot.it = FALSE)
    }
    else {
      n.trees = length(object$train.error)
    }
    cat(paste("Using", n.trees, "trees...\n"))
    gbm::predict.gbm(object, newdata, n.trees, type, single.tree, ...)
  }
}

logit2prob = function(x){
  exp(x) / (1 + exp(x))
}

prob2odds = function(x){
  x / (1 - x)
}

threshold_compare = function(thresh, dataframe, model){
  pred = dataframe %>%
         add_predictions(model) %>%
         mutate(pred = logit2prob(pred),
         pred = as.numeric(pred > thresh))
}
```

# Part 1: Joe Biden (redux times two)

We read in the data:

```{r read_data_biden}

df = read.csv('data/biden.csv')
```

Next, we split the dataset into training and validation sets in a ratio of 7:3, then estimate a regression tree:

```{r first_tree}

set.seed(1234)

biden_split7030 = resample_partition(df, c(test = 0.3, train = 0.7))
biden_train70 = biden_split7030$train %>%
                tbl_df()
biden_test30 = biden_split7030$test %>%
               tbl_df()

# estimate model
biden_tree1 = tree(biden ~ female + age + educ + dem + rep, data = biden_train70)
mse_test30 = mse(biden_tree1,biden_test30)
```

We evaluate the model with the testing data and find that the mean squared error is `r mse_test30`.  Next, we plot the tree:

```{r plot_biden_tree1, echo=FALSE}

plot(biden_tree1, col='darkturquoise', lwd=2.5)
title("Regression Tree for Warmth Toward Joe Biden (2008)\n", sub = "Validation Dataset")
text(biden_tree1, col='deeppink')
```

The model shows that identifying as a Democrat is the strongest predictor of feelings of warmth toward Vice President Biden, and identifying as a Republican is the second-strongest predictor.  Together, these splits indicate that party affiliation (whether Democratic, Republican, or neither) is the most important factor when it comes to predicting an individual subject's feelings of warmth toward Mr. Biden.

We now fit another, more complex tree model:
```{r biden_tree_2}
biden_tree2 = tree(biden ~ female + age + educ + dem + rep, data = biden_train70,
                   control = tree.control(nobs = nrow(biden_train70), mindev = .0000001))
```

We prune the tree 49 different times, increasing the number of leaves from 2 to 50 and storing the MSE for each pruned tree:

```{r biden_prune_50_rounds}

rounds = 50

mse_list_biden_50 = vector("numeric", rounds - 1)
leaf_list_biden_50 = vector("numeric", rounds - 1)

set.seed(1234)

for(i in 2:rounds) {
    biden_mod = prune.tree(biden_tree2, best=i)

    mse_val = mse(biden_mod,biden_test30)
    mse_list_biden_50[[i-1]] = mse_val
    leaf_list_biden_50[[i-1]] = i
}

mse_df_biden_50 = as.data.frame(mse_list_biden_50)
mse_df_biden_50$branches = leaf_list_biden_50
```

We plot the MSE for each tree vs. the number of leaves:

```{r plot biden_50_mse, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(mse_df_biden_50, aes(branches, mse_list_biden_50)) +
       geom_line(color='darkturquoise',size=1) +
       labs(title = "Comparing Regression Trees for Warmth Toward Joe Biden",
       subtitle = "Using Validation Set",
       x = "Number of Leaves",
       y = "Mean Squared Error") + 
       theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
             panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1))
```

We can clearly see that the lowest MSE is `r min(mse_df_biden_50$mse_list_biden_50)` for a tree with 11 leaves.  We plot that tree:

```{r plotting_optimal_biden_tree}

biden_pruned11 <- prune.tree(biden_tree2, best=11)
mse_test30_2 = mse(biden_pruned11,biden_test30)

plot(biden_pruned11, col='darkturquoise', lwd=2.5)
title("Regression Tree (Best 11) for Warmth Toward Joe Biden (2008)\n", sub = "Validation Dataset")
text(biden_pruned11, col='deeppink')
```

The tree indicates that for Democrats, age is the next most important variable and education after that, but that gender is not important.  For unaffiliated voters, gender is important; for women, education and age are both important, but not so for men.  Among Republican voters, age is important within the whole group and education is important, but only for voters between ages 44 and 47; gender is not an important predictor of feelings of warmth toward Joe Biden among Republican voters.

Pruning the tree reduces the MSE from `r mse_test30` to `r mse_test30_2`.

We use the bagging approach to analyze this data, computing 500 bootstrapped trees using the training data and testing the resulting model with the validation set:

```{r bagging_biden}

set.seed(1234)

biden_bag_data_train = biden_train70 %>%
                       rename() %>%
                       mutate_each(funs(as.factor(.)), dem, rep) %>%
                       na.omit

biden_bag_data_test = biden_test30 %>%
                      rename() %>%
                      mutate_each(funs(as.factor(.)), dem, rep) %>%
                      na.omit

(bag_biden <- randomForest(biden ~ ., data = biden_bag_data_train, mtry = 5, ntree = 500, importance=TRUE))

mse_bag_biden = mse(bag_biden, biden_bag_data_test)

bag_biden_importance = as.data.frame(importance(bag_biden))
```

Using the validation data, the model returns a test MSE of `r mse_bag_biden`, which is considerably higher than the MSE found when pruning the tree earlier (`r mse_test30_2`).

Next, we review variable importance measures:

```{r biden_bag_importance, echo=FALSE}

ggplot(bag_biden_importance, mapping=aes(x=rownames(bag_biden_importance), y=IncNodePurity)) +
       geom_bar(stat="identity", aes(fill=IncNodePurity)) +
       labs(title = "Average Increased Node Purity Across 500 Bagged Regression Trees",
       subtitle = "Predicted Warmth Toward Joe Biden (2008)",
       x = "Variable",
       y = "Mean Increased Node Purity") + 
       theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
             panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1), legend.position = 'none') 
```

The variable importance plot shows that age and Democrat are the two most important variables as these yield the greatest average decreases (of approximately 225,000 and 150,000, respectively) in node impurity across 500 bagged regression trees.  Despite the higher test MSE, the bagged tree model is likely a better model than the pruned tree above because the bagged model uses bootstrapping to create 500 different training sets, whereas the pruned tree above uses only a single training set.  The bagged model averages the variance across the bootstrapped trees, which, together, suggest that age and Democrat are the most important variables while gender is the least important.  It is worth noting, however, that the bagged model only accounts for 9.49% of the variance in feelings of warmth toward Joe Biden.

Next, we estimate a random forest model with 500 trees:

```{r biden_random_forest}

set.seed(1234)

m = floor(sqrt(5))

(rf_biden = randomForest(biden ~ ., data = biden_bag_data_train, mtry = m, ntree = 500))
mse_rf_biden = mse(rf_biden, biden_bag_data_test)

```

The random forest model returns a test MSE of `r mse_rf_biden`, which is much lower than the one returned by bagging (`r mse_bag_biden`).  Furthermore, the random forest model explains a greater proportion of variance (25.9%) than the bagged model does (9.49%).  Still, with the variane explained at only a quarter, this suggests that there are likely unobserved and unknown variables that have a notable effect on feelings of warmth for Joe Biden.

The notable decrease in MSE is attributable to the effect of limiting the variables available every split to only $m$ (`r m`) randomly-selected predictors.  This means that the trees in the random forest model will be uncorrelated to each other, the variance for the final model will be lower, and the test MSE will be lower.

We plot the importance of the predictors:

```{r plot_rf_importance, echo=FALSE}
rf_biden_importance = as.data.frame(importance(rf_biden))

ggplot(rf_biden_importance, mapping=aes(x=rownames(rf_biden_importance), y=IncNodePurity)) +
       geom_bar(stat="identity", aes(fill=IncNodePurity)) + 
       labs(title = "Average Increased Node Purity Across 500 Random Forest Regression Trees",
       subtitle = "Predicted Warmth Toward Joe Biden (2008)",
       x = "Variable",
       y = "Mean Increased Node Purity") + 
       theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
             panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1), legend.position = 'none') 
```

The random forest model estimates that Democrat is the sinlge most important predictor of feelings toward Joe Biden and that Republican is next in line; these have mean increased node purity of approximately 110,000 and 70,000, respectively.  As was the case with the bagging model, gender is the least important predictor.

Finally, we estimate three boosting models, each of different depths and with 10,000 trees:

```{r biden_boosting_models}

set.seed(1234)
biden_models = list("boosting_depth1" = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train,
                                            n.trees = 10000, interaction.depth = 1),
                    "boosting_depth2" = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train,
                                            n.trees = 10000, interaction.depth = 2),
                    "boosting_depth4" = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train,
                                            n.trees = 10000, interaction.depth = 4))
```

For each depth, we find the optimal number of iterations:

```{r biden_boosting_iterations}

set.seed(1234)
data_frame(depth = c(1, 2, 4),
           model = biden_models[c("boosting_depth1", "boosting_depth2", "boosting_depth4")],
           optimal = map_dbl(model, gbm.perf, plot.it = FALSE)) %>%
           select(-model) %>%
           knitr::kable(caption = "Optimal number of boosting iterations",
                        col.names = c("Depth", "Optimal number of iterations"))
```

Now we estimate the boosting models with the optimal number of treesh for each depth:

```{r biden_boosting_optimal_trees}

set.seed(1234)

biden_boost1 = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train, n.trees = 3302, interaction.depth = 1)

biden_boost2 = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train, n.trees = 2700, interaction.depth = 2)

biden_boost4 = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train, n.trees = 2094, interaction.depth = 4)

mse_boost1_biden = mse(biden_boost1,biden_bag_data_test)
mse_boost2_biden = mse(biden_boost2,biden_bag_data_test)
mse_boost4_biden = mse(biden_boost4,biden_bag_data_test)
```

The boosting model with a depth of 1 has a test MSE of `r mse_boost1_biden`; for the model with a depth of 2, it is `r mse_boost2_biden` and for the model with a depth of 4 it is `r mse_boost4_biden`.  This indicates that the boosting approach yields the lowest MSE for trees with a single split compared to those with two or four splits.

Next, we increase the value of the $\lambda$ from the default of .001 to .1:

```{r stuff3}

set.seed(1234)

boost1_biden_lambda = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train,
                          n.trees = 3302, interaction.depth = 1, shrinkage=0.1)

boost2_biden_lambda = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train,
                          n.trees = 2700, interaction.depth = 2, shrinkage=0.1)

boost4_biden_lambda = gbm(as.numeric(biden) - 1 ~ ., data = biden_bag_data_train,
                          n.trees = 2094, interaction.depth = 4, shrinkage=0.1)

mse_boost1_biden_lambda = mse(boost1_biden_lambda,biden_bag_data_test)
mse_boost2_biden_lambda = mse(boost2_biden_lambda,biden_bag_data_test)
mse_boost4_biden_lambda = mse(boost4_biden_lambda,biden_bag_data_test)
```

The test MSE for single-split trees has increased from `r mse_boost1_biden` to `r mse_boost1_biden_lambda`; for trees with a depth of two, it has increased from `r mse_boost2_biden` to `r mse_boost2_biden_lambda` and for trees of depth 4, it has increased from `r mse_boost4_biden` to `r mse_boost4_biden_lambda`.  This suggests that increasing the step size leads to the model learning faster but not as well.  Ideally, the next step would be to try different values of $\lambda$ and determine which yields the lowest MSE for each depth.


# Part 2: Modeling voter turnout

An important question in American politics is why do some people participate in the political process, while others do not? Participation has a direct impact on outcomes -- if you fail to participate in politics, the government and political officials are less likely to respond to your concerns. Typical explanations focus on a resource model of participation -- individuals with greater resources, such as time, money, and civic skills, are more likely to participate in politics. One area of importance is understanding voter turnout, or why people participate in elections. Using the resource model of participation as a guide, we can develop several expectations. First, women, who more frequently are the primary caregiver for children and earn a lower income, are less likely to participate in elections than men. Second, older Americans, who typically have more time and higher incomes available to participate in politics, should be more likely to participate in elections than younger Americans. Finally, individuals with more years of education, who are generally more interested in politics and understand the value and benefits of participating in politics, are more likely to participate in elections than individuals with fewer years of education.

While these explanations have been repeatedly tested by political scientists, an emerging theory assesses an individual's mental health and its effect on political participation.^[[Ojeda, C. (2015). Depression and political participation. *Social Science Quarterly*, 96(5), 1226-1243.](http://onlinelibrary.wiley.com.proxy.uchicago.edu/doi/10.1111/ssqu.12173/abstract)] Depression increases individuals' feelings of hopelessness and political efficacy, so depressed individuals will have less desire to participate in politics. More importantly to our resource model of participation, individuals with depression suffer physical ailments such as a lack of energy, headaches, and muscle soreness which drain an individual's energy and requires time and money to receive treatment. For these reasons, we should expect that individuals with depression are less likely to participate in election than those without symptoms of depression.

The 1998 General Social Survey included several questions about the respondent's mental health. `mental_health.csv` reports several important variables from this survey.

* `vote96` - 1 if the respondent voted in the 1996 presidential election, 0 otherwise
* `mhealth_sum` - index variable which assesses the respondent's mental health, ranging from 0 (an individual with no depressed mood) to 9 (an individual with the most severe depressed mood)^[The variable is an index which combines responses to four different questions: "In the past 30
days, how often did you feel: 1) so sad nothing could cheer you up, 2) hopeless, 3) that everything was an effort, and 4) worthless?" Valid responses are none of the time, a little of the time, some of the time, most of the time, and all of the time.]
* `age` - age of the respondent
* `educ` - Number of years of formal education completed by the respondent
* `black` - 1 if the respondent is black, 0 otherwise
* `female` - 1 if the respondent is female, 0 if male
* `married` - 1 if the respondent is currently married, 0 otherwise
* `inc10` - Family income, in \$10,000s

```
(mh <- read_csv("data/mental_health.csv") %>%
  mutate_each(funs(as.factor(.)), vote96, black, female, married) %>%
  na.omit)
  ```






1. Use cross-validation techniques and standard measures of model fit (e.g. test error rate, PRE, ROC curves/AUC) to compare and evaluate at least five tree-based models of voter turnout. Select the best model and interpret the results using whatever methods you see fit (graphs, tables, model fit statistics, predictions for hypothetical observations, etc.)
1. Use cross-validation techniques and standard measures of model fit (e.g. test error rate, PRE, ROC curves/AUC) to compare and evaluate at least five SVM models of voter turnout. Select the best model and interpret the results using whatever methods you see fit (graphs, tables, model fit statistics, predictions for hypothetical observations, etc.)

# Part 3: OJ Simpson



1. What is the relationship between race and belief of OJ Simpson's guilt? Develop a robust statistical learning model and use this model to explain the impact of an individual's race on their beliefs about OJ Simpson's guilt.

1. How can you predict whether individuals believe OJ Simpson to be guilty of these murders? Develop a robust statistical learning model to predict whether individuals believe OJ Simpson to be either probably guilty or probably not guilty and demonstrate the effectiveness of this model using methods we have discussed in class.

Select methods that are appropriate for each question and **justify the use of these methods**.

We begin by reading in the data and dropping those observations that are missing data in the response variable, guilt (for which we also make into a factor and assign labels):

```{r read_data_biden}

oj = read.csv('data/simpson.csv')
oj = oj[(!is.na(oj$guilt)), ]
oj$Opinion = factor(oj$guilt, levels = c(0,1), labels = c("Probably not guilty", "Probably guilty"))
```








```{r oj_histogram_black, echo=FALSE}
ggplot(oj, aes(x=black, fill=Opinion)) + geom_bar(position = "dodge") + 
       ylab("Frequency count of respondents") +
       xlab("Race") +
       ggtitle("Opinion of Simpson Guilt Based on Race") +
       theme(plot.title = element_text(hjust = 0.5),
       panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(),
       panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1)) +
       scale_x_continuous(breaks = c(0,1), labels = c("Not Black", "Black"))
```





```{r oj_basic_logistic_regression}

logit_oj_black = glm(guilt ~ black, family = binomial, data=oj)
summary(logit_oj_black)

mse_logit_oj_black = mse(logit_oj_black, oj)
```




```{r bv_gen_dfs_and_vars}

int = tidy(logit_oj_black)[1,2]
coeff = tidy(logit_oj_black)[2,2]

oj_black_pred = oj %>%
                add_predictions(logit_oj_black) %>%
                mutate(prob = logit2prob(pred)) %>%
                mutate(odds = prob2odds(prob))

oj_black_accuracy = oj %>%
                    add_predictions(logit_oj_black) %>%
                    mutate(pred = logit2prob(pred),
                    pred = as.numeric(pred > .5))
```



```{r, ar_auc}

ar = mean(oj_black_accuracy$guilt == oj_black_accuracy$pred, na.rm = TRUE)

uc = median(oj$guilt)

cm.5_oj_black = confusionMatrix(oj_black_accuracy$pred, oj_black_accuracy$guilt,
                                dnn = c("Prediction", "Actual"), positive = '1')

cm.5_table = cm.5_oj_black$table

actlpos = cm.5_table[1,2] + cm.5_table[2,2]
predposcrrct = cm.5_table[2,2]

actlneg = cm.5_table[1,1] + cm.5_table[2,1]
prednegcrrct = cm.5_table[1,1]

tpr.notes =  predposcrrct / actlpos
tnr.notes =  prednegcrrct / actlneg

tpr.cm.5 = sum(cm.5_oj_black$byClass[1])
tnr.cm.5 = sum(cm.5_oj_black$byClass[2])


```
















Next, we split the dataset into training and validation sets in a ratio of 7:3:

```{r first_tree}

set.seed(1234)

oj_split7030 = resample_partition(oj, c(test = 0.3, train = 0.7))
oj_train70 = oj_split7030$train %>%
             tbl_df()
oj_test30 = oj_split7030$test %>%
            tbl_df()
```



```{r oj_random_forest}

set.seed(1234)

oj_data_train = oj_train70 %>%
                select(-guilt) %>%
                mutate_each(funs(as.factor(.)), dem, rep) %>%
                na.omit

oj_data_test = oj_test30 %>%
               select(-guilt) %>%
               mutate_each(funs(as.factor(.)), dem, rep) %>%
               na.omit

m_oj = floor(sqrt(9))

(rf_oj = randomForest(Opinion ~ ., data = oj_data_train, mtry = m_oj, ntree = 500))
```




```{r plot_rf_importance_oj, echo=FALSE}
rf_oj_importance = as.data.frame(importance(rf_oj))

ggplot(rf_oj_importance, mapping=aes(x=rownames(rf_oj_importance), y=MeanDecreaseGini)) +
       geom_bar(stat="identity", aes(fill=MeanDecreaseGini)) + 
       labs(title = "Mean Decrease in Gini Index Across 500 Random Forest Regression Trees",
       subtitle = "Predicted Opinion of Simpson Guilt",
       x = "Variable",
       y = "Mean Decrease in Gini Index") + 
       theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
             panel.border = element_rect(linetype = "solid", color = "grey70", fill=NA, size=1.1), legend.position = 'none') 
```