---
title: "PS6"
author: "Xinzhu Sun"
date: "2/17/2017"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```
````{r library}
library(tidyverse)
library(modelr)
library(broom)
library(dplyr)
library(ggplot2)
library(readr)
library(forcats)
library(pROC)
options(na.action = na.warn)
set.seed(1234)
theme_set(theme_minimal())
```

## Part 1： Modoling voter turnout
## Describe the data
1.Plot a histogram of voter turnout.
```{r histogram}
mental_health <- read_csv("mental_health.csv")
health_nona <- mental_health[!is.na(mental_health$vote96),]
ggplot(mental_health, mapping = aes(x = vote96)) +
  geom_histogram() +
  labs(title = "Distribution of voter turnout",
       x = "voter turnout",
       y = "Frequency count of votoer turnout")
prob_vote = sum(health_nona$vote96 == 1)/nrow(health_nona)
```
The unconditional probability of a given individual truning out to vote is:  0.6823574. 

2.Generate a scatterplot of the relationship between mental health and observed voter turnout and overlay a linear smoothing line.
```{r scatterplot}
ggplot(mental_health, aes(mhealth_sum,vote96)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(y = 'voter turnout',
       x = 'index variable of mental health',
       title= 'Scatterplot of the relationship between mental health and observed voter turnout')
```

This scatterplot tells us that whether voter turnout or not is negatively corelated with mental health.

The ploblem of this linear smoothing line is obviously, it doesn't fit the scatters well. First, the only possible values for voter turnout are 0 and 1. Yet the linear regression model gives us predicted values between 0 and 1, such as 0.75 and 0.25. Second, because the line is linear and continuous, it extends infinitely in both directions of mental health.

## Basic model
```{r logistic regression}
voter_health <- glm(vote96 ~ mhealth_sum, data = mental_health, family = binomial)
summary(voter_health)
```

1.Yes, the relationship between voter turnout and meantal health is satistically and substantively significant. The esitmated parameter is about -0.14348 and the significance level is on 0.001.

2. When the evaluation on mental health index increases by one unit, the log odds of voter voting against not voting decreases by -0.14348. 
```{r log-odds}
voter_health_pred <- mental_health %>%
  add_predictions(voter_health)

ggplot(voter_health_pred, aes(pred, mhealth_sum)) +
  geom_line(size = 1)+
  labs(
    x = 'index variable of mental health',
    y = 'log-odds of voter turnout',
    title = 'relationship between mental health and the log-odds of voter turnout'
  )
```

3.The estimator on odds can be interpreted as percent change. When the evaluation on mental health increases by one unit, the odds of voter voting against not voting decreases by -14.348 percent(%).  
```{r odds}
voter_health_pred <- voter_health_pred %>%
  mutate(odds = exp(pred))

ggplot(voter_health_pred, aes(odds, mhealth_sum)) +
  geom_line(szie = 1)+
  labs(
    x = 'index variable of mental health',
    y = 'odds of voter turnout',
    title = 'relationship between mental health and the odds of voter turnout'
  )
```

4.The interpretation of the estimator from the perspective of probablity is not clear. Since the first difference typically depend on the initial age.
```{r probabilities}
logit2prob<-function(x){
  exp(x)/(1 + exp(x))
}
voter_health_pred <- voter_health_pred %>%
  mutate(prob = logit2prob(pred))

ggplot(voter_health_pred, aes(odds, mhealth_sum)) +
  geom_line(szie = 1)+
  labs(
    x = 'index variable of mental health',
    y = 'probability of voter turnout',
    title = 'relationship between mental health and the probability of voter turnout'
  )
fd_pred <- data.frame(mhealth_sum = c(1, 2, 5, 6))%>%
  add_predictions(voter_health)%>%
  mutate(prob = logit2prob(pred))
fd1 = fd_pred[2, ]$prob - fd_pred[1, ]$prob
print(paste('first difference for an increase in mental health index from 1 to 2 is ', fd1))
fd2 = fd_pred[4, ]$prob - fd_pred[3, ]$prob
print(paste('first difference for an increase in mental health index from 5 to 6 is ', fd2))
```

The first difference for an increase in the mental health index from 1 to 2 is -0.0292; from 5 to 6 is -0.0348.

```{r accuracy rate, PRE, AUC}
clean_data = mental_health[!is.na(mental_health$vote96) & !is.na(mental_health$mhealth_sum),]
mental_accuracy <- clean_data%>%
  add_predictions(voter_health)%>%
  mutate(prob = logit2prob(pred),
         pred = as.numeric(prob >.5))
mean(mental_accuracy$pred == mental_accuracy$vote96, na.rm = TRUE)

null_err <- nrow(clean_data[clean_data$vote96 == 0,])
model_err <- sum(mental_accuracy$pred != mental_accuracy$vote96)
pre <- (null_err - model_err)/null_err
pre

auc_m <- auc(mental_accuracy$vote96, mental_accuracy$prob)
auc_m
```

5.The accuracy rate is 0.6778. The proportional reduction in error is 0.01616628. The AUC is 0.6243. I don't think it's a good model. The proportional reduction in error is only around 1.6%, which is very small thus the model doesn't explain the binary choice of voting well.

## Multiple variable model
1.The three componets of the GLM:
  Probability distribution(random component)
  $$Pr(Y_i = y_i | \pi_i) = (\pi_i)^{y_i}(1 - \pi_i)^{1 - y_i}$$
  $\pi_i$ is the population 'mean' we want to model.
  
  Linear predictor: 
  $$\begin{aligned}
  \eta_i = &\beta_0 + \beta_1 mhealth_sum_i + \beta_2 age_i + \beta_3 educ_i + \\
  &\beta_4 black_i + \beta_5 black_i + \beta_6 female_i + \beta_7 married_i + \beta_8 inc10_i
  \end{aligned}$$
  
  Link function:
  $$\pi_i = g(\eta_i) = \frac{e^{\eta_i}}{1 + e^{\eta_i}}$$
2.The estimated multiple variable model is reported as following:
```{r estimate the model}
multiple_vote <- glm(vote96 ~ mhealth_sum + age + educ + black + female + married + inc10, data = clean_data, family = binomial)
summary(multiple_vote)
```

```{r interpret the results}
clean_data <- na.omit(mental_health)
mental_accuracy <- clean_data%>%
  add_predictions(multiple_vote)%>%
  mutate(prob = logit2prob(pred),
         pred = as.numeric(prob >.5))

null_err <- nrow(clean_data[clean_data$vote96 == 0,])
model_err <- sum(mental_accuracy$pred != mental_accuracy$vote96)
pre <- (null_err - model_err)/null_err
pre
```
3.Overall, the preformanceof this model improves significantly comparing to the basic model.We can see that the PRE is 0.1481481 now while in the basic model, PRE is only 0.01616628.

Among all independent variables, the mental health index, age, education and income turn out to be significant on the significance level of 0.01. Mental health index has a negative relationship with the voter turnout. On average, one level increase in the mental health index will reduce the odds by 8.9102%. On the other hand, age, education and income all have positive effect on voter turnout. Specifically, one year increase in age will increase the odds of voting by 4.2534%; one year increase in years of educatoin will on average increase the odds of voting by 22.8686%; and every ten thousand dollar increase in income will on average increase the odds of voting by 6.9614%.  
  
Marriage status is significant on the 0.1 significance level and is very close to 0.05. If one is married, then his or her odds of voting will increase by 29.6915% in general. However, whether a person is black, and gender is statistically insignificant. 

We can plot the predicted probablity of voter turnout against the index of mental health and divide the data into four groups: married black, married non-black, unmarried black and unmarried non-black.

```{r plots}  
mental_accuracy_pred <- clean_data%>%
  data_grid(mhealth_sum, black, married)%>%
  cbind(data_frame(age = mean(clean_data$age),
                   educ = mean(clean_data$educ),
                   female = median(clean_data$female),
                   inc10 = mean(clean_data$inc10)
                   ))%>%
  add_predictions(multiple_vote)%>%
  mutate(prob = logit2prob(pred))

ggplot(mental_accuracy_pred, aes(x = mhealth_sum, y = prob))+
  geom_line(size = 1, aes(group = interaction(black, married), color = interaction(black, married)))+
  scale_color_discrete(name = 'Group',
    labels = c('unmarried non-black', 'unmarried black', 'married non-black', 'married black'))+
  labs(x = 'index variable of mental health',
       y = 'probability of voter turnout',
       title = 'relationship between mental health and the probability of voter turnout(Black X Marriage)')
```  

We can from the graph, married people and black people both have higher probablity of voting. Though we see an increase of almost 0.1 in probablity, we still can't get any conclusion. On the one hand these variables are statistically insignificant on 0.05 level, on the other hand a 0.1 incrase in voting probablity is not a negligible effect. Serveral factors can casue this problem. The most probable one is multicolinearity with other variables. For example, marriage status can be closely correlated with age and income. 

## Part2: Modelling tv consumption
## Estimate a regression model
1.The three componets of the GLM:
  Probability distribution(random component)
  $$Pr(Y_i = y_i | \mu_i) = \frac{\mu^k e^{-y_i}}{y_i!}$$
  $\pi_i$ is the population 'mean' we want to model;
  
  Linear predictor: 
  $$\begin{aligned}
  \eta_i = &\beta_0 + \beta_1 age_i + \beta_2 children_i + \beta_3 education_i + \beta_4 female_i + \beta_5 grass_i + \\
  &\beta_6 hrsrelax_i + \beta_7 black_i + \beta_8 social\_connect_i + \beta_9 voted04_i + \\
  &\beta_{10} xmovie_i + \beta_{11} zodiac_i
  \end{aligned}
  $$
  
  Link function:
  $$log(\mu_i) = \eta_i$$

2.The estimated poisson model is reported as following:
```{r possion}
tv_data <- read_csv('gss2006.csv')
multiple_tv <- glm(tvhours ~ age + childs + educ + female + grass + hrsrelax + black + social_connect + voted04 + xmovie + zodiac, data = tv_data, family = poisson)
summary(multiple_tv)
```

3.We can see from the regression table, years of edcuation, hours of relax per day and whether people are black are significant on 0.005 level. Specifically, 1 year incrase in education on average cause 2.9077% decrease in hours of watching TV; one hour incrase in hours of relax per day on average causes 4.6556% increase in hours of watching TV; holding all other constant, black people on average wactch 46.2% more hours of TV,This is a farily large incrase. 

Thus the results tell us,as hours of relax increases, number of hours of watching television also increases. But as hours of relaxation increases, watching TV is not the only way of entertainment. Maybe because after hours of relax increase to a certain amount, people invest more time to other types of entertainment. There should be a non-linear relationship between hours of wathcing TV and hours of relax. Let's the following regression adding the square of hours of relax:
```{r futher}
tv_data <- read_csv('gss2006.csv')
multiple_tv_2 <- glm(tvhours ~ age + childs + educ + female + grass + I(hrsrelax^2) + black + social_connect + voted04 + xmovie + zodiac, data = tv_data, family = poisson)
summary(multiple_tv_2)
```

We can see that the square of hours of relax is negatively significant. This suggests there is a firstly incrasing and then decreasing relationship between hours of relax and hours of watching TV. We then plot the predicted counts against hours of relax. In this case, we take all other predictors as their median value (except for zodiac, I took Aries as the predictor value):
```{r plot}
tv_pred <- tv_data[!is.na(tv_data$hrsrelax), c('hrsrelax', 'tvhours')]%>%
  cbind(data_frame(
  age = mean(tv_data$age, na.rm = TRUE),
  childs = mean(tv_data$childs, na.rm = TRUE),
  educ = mean(tv_data$educ, na.rm = TRUE),
  female = median(tv_data$female, na.rm = TRUE),
  grass = median(tv_data$grass, na.rm = TRUE),
  black = median(tv_data$black, na.rm = TRUE),
  social_connect = mean(tv_data$social_connect, na.rm = TRUE),
  voted04 = median(tv_data$voted04, na.rm = TRUE),
  xmovie =median(tv_data$xmovie, na.rm = TRUE),
  zodiac = 'Aries'
  )) %>%
  add_predictions(multiple_tv)%>%
  mutate(count = exp(pred))
ggplot(tv_pred, aes(x = hrsrelax, y = count))+
  geom_line()+
  geom_point(aes(y = tvhours), alpha = 0.5)+
  labs(
    x = 'Hours of Relax',
    y = 'Predicted Hours of Wathcing TV',
    title = 'Predcited Hours of Watching TV vs. Hours of Relax'
  )
```

As for other predictors, they are not significant in 0.1 level. But notice, zodiac is not related to hours of watching TV. This may be a supporting evidence that zodiac is just a relfect of random month to be born in.

Interpretting the other aspects of our model, we can realize that age, gender, number of children and party affilication all have no statistically significant influcence on TV consumption. One could argue that hours of relaxation and years of education could reflect some of these variables, but they are also intuitively directly related to tv consumption.

And also, more well-educated one person is, less hours TV he or she watches. This may come as a result of having more work to do for their job or they have other ways of entertainment. 

And black people tend to watch more TV. Specifically, let's see one more plot of the impact of black or not on prdicted hours of watching TV per day, ignore unrelavent variables:
```{r futher plot}
poisson_1 <- glm(tvhours ~ educ + hrsrelax + black, data = tv_data, family = quasipoisson)
tv_poisson_1_grid <- tv_data %>%
  data_grid(educ, hrsrelax, black) %>%
  add_predictions(poisson_1) %>%
  mutate(count_pred = exp(pred))

ggplot(tv_poisson_1_grid, aes(hrsrelax, count_pred, group = factor(black), color = factor(black))) +
  geom_line() +
  scale_color_discrete(name = "Black (is black = 1)") +
  labs(title = "Predicted hours of TV watched per day",
       subtitle = "by hours has to relax, black or not, and years of education",
       x = "Hours has to relax per day",
       y = "Predicted hours of TV watched per day")
```

As in the graph, we can observe the interactive relationships between the independent variables in the model. As hours has to relax goes up, the distance between the red and green lines are larger, indicating the larger difference between black and non-black people at higher levels of relaxing hours, regarding the time spending on watching TV. Also, the lines have larger fluctuation at higher levels of relaxing hours, which indicates that when relaxing hour goes up, the effect of education on tvhours also goes up. Lastly, the fluctuation is more drastical for black people than non-black people, that infers the education's effect on black people is more than on non-black people regarding the TV consumption. This result on its surface doesn't make sense. There could be other socio-economic factors that we don't take into account, for example, income, whether in a food stamp program, or whether on a social welfare program, etc.



